\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?\newline Evidence from self-paced reading and event-related potentials using German

### 2023-10-10 paper/main.tex
minor edit

It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the pre-critical and critical regions. The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power. In order to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region, we inserted two adverbs. This way, the critical region will be less affected by the clause-boundary slowdown. This design also has the advantage that any potential effects seen at the pre-critical region (the adverb preceding the verb) are less likely to be due to spill-over; it is important to look at the pre-critical region as previous work with this design found effects at the pre-critical region (we revisit this detail in the General discussion).

@@

The SPR study, hereafter Experiment 1a, was initially designed so that participants  (N=160) read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120). As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: the effects observed in the first session disappeared. The adaptation effect is theoretically interesting per se \parencite{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation. For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=614) were either shown items 1-60 or items 61-120 (but not both). In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
As mentioned above, because the ERP method requires a large number of items, all the data come from participants who completed both sessions.

>>>>

\paragraph{Regions of interest in the design}

Although our principal focus was on the reading times at the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region. A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would cause two verbs to appear one after another ( \textit{erzählt hatte, trank}, literally `told drank'). This is because in German, the verb is clause-final in embedded clauses. Such verb sequences are inherently difficult to process \parencite{VSLK11,bach1986crossed}. T

For these reasons, we added two adverbs before the critical verb. These intervening regions had a further advantage: 
It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region. The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power. 
Inserting two adverbs before the critical verb thus served to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region. This way, the critical region will be less affected by the clause-boundary slowdown. This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.

Thus, our focus in the present study is on the processing difficulty observed at the pre-critical, critical (verb), and post-critical regions.

@@

The SPR study, hereafter Experiment 1a, was initially designed so that participants  (N=160) read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120). The variation between participants in the gap is due to the fact that we could not control when the participants would do the second session; however, the mean and median gaps were  about 11 days, with the first and third quartiles being 9 and 12 days.

As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: the effects observed in the first session disappeared. The adaptation effect is theoretically interesting per se \parencite{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation. For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=614) were either shown items 1-60 or items 61-120 (but not both). In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
As mentioned above, because the ERP method requires a large number of items, all the data come from participants who completed both sessions.
